Predicting FAANG Stock Prices: A Data-Driven Approach Using Machine Learning Models

### 2024-12-03 paper/paper.qmd
committing changes to results section in paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

@@

@@

# Results

Our results are summarized in @tbl-modelresults.

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

>>>>

---
title: "Predicting FAANG Stock Prices: A Data-Driven Approach Using Machine Learning Models"
subtitle: "My subtitle if needed"
author: 
  - Jimin (Jamie) Lee
thanks: "Code and data are available at: https://github.com/jamiejiminlee/FAANG-Stock-Forecast.git."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(arrow)
library(kableExtra)
library(xgboost)
library(tidyr)
library(zoo)
library(scales)
library(reshape2)

analysis_data <- read_parquet("/Users/jamielee/TechStockForecast/data/02-analysis_data/analysis_data.parquet")
```

@@

library(xgboost)

nov_model <- readRDS("/Users/jamielee/TechStockForecast/models/nov_model.rds")
dec_model <- readRDS("/Users/jamielee/TechStockForecast/models/dec_model.rds")

```

@@

# Results
```{r}
# Filter November data for prediction
nov_test <- analysis_data %>%
  filter(as.Date(date) >= as.Date("2024-11-01") & as.Date(date) <= as.Date("2024-11-30"))

# Prepare testing features for November
X_test_nov <- nov_test %>%
  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
  as.matrix()

# Predict November prices
dtest_nov <- xgb.DMatrix(data = X_test_nov)
nov_predictions <- predict(nov_model, dtest_nov)

# Combine predictions with actual November data
nov_results <- nov_test %>%
  mutate(predicted_adjusted = nov_predictions) %>%
  select(date, symbol, adjusted, predicted_adjusted)

```

```{r}
# Filter December data for prediction
dec_test <- analysis_data %>%
  filter(as.Date(date) >= as.Date("2024-12-01") & as.Date(date) <= as.Date("2024-12-31"))

# Prepare testing features for December
X_test_dec <- dec_test %>%
  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
  as.matrix()

# Predict December prices
dtest_dec <- xgb.DMatrix(data = X_test_dec)
dec_predictions <- predict(dec_model, dtest_dec)

# Combine predictions with actual December data
dec_results <- dec_test %>%
  mutate(predicted_adjusted = dec_predictions) %>%
  select(date, symbol, predicted_adjusted)

```

## Actual vs Predicted for November 2024
```{r}

# Plot actual vs predicted prices for November 2024
ggplot(nov_results, aes(x = as.Date(date))) +
  geom_line(aes(y = adjusted, color = "Actual"), size = 1) +  # Actual prices
  geom_line(aes(y = predicted_adjusted, color = "Predicted"), linetype = "dashed", size = 1) +  # Predicted prices
  facet_wrap(~symbol, scales = "free_y") +  # Separate panels for each stock
  labs(
    title = "Actual vs Predicted Stock Prices for November 2024",
    x = "Date",
    y = "Stock Price",
    color = "Legend"
  ) +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) +  # Define colors
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Heatmap of Prediction Errors (November 2024)
```{r}
# Add the `error` column to `nov_results`
nov_results <- nov_results %>%
  mutate(error = abs(adjusted - predicted_adjusted))  # Calculate absolute error

# Prepare data for heatmap
nov_heatmap_data <- nov_results %>%
  select(date, symbol, error) %>%
  spread(symbol, error)

# Plot heatmap of prediction errors
ggplot(melt(nov_heatmap_data, id.vars = "date"), aes(x = date, y = variable, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Heatmap of Prediction Errors for November 2024",
    x = "Date",
    y = "Stock Symbol",
    fill = "Error"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 12)
  )
```

## Average Predicted Stock Price - November
```{r}
# Calculate average predicted prices for November
nov_avg_prices <- nov_results %>%
  group_by(symbol) %>%
  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE))

# Plot average predicted prices
ggplot(nov_avg_prices, aes(x = symbol, y = avg_predicted_price, fill = symbol)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    title = "Average Predicted Stock Prices for November 2024",
    x = "Stock Symbol",
    y = "Average Predicted Price",
    fill = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_brewer(palette = "Set2")

```

## Predicted Stock Price Trends - December
```{r}
ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
  geom_line(size = 1) +
  labs(
    title = "Predicted Stock Price Trends for December 2024",
    x = "Date",
    y = "Predicted Stock Price",
    color = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Average Stock Prices Bar Chart - December
```{r}
# Calculate average predicted prices for December
dec_avg_prices <- dec_results %>%
  group_by(symbol) %>%
  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE))

# Plot average predicted prices
ggplot(dec_avg_prices, aes(x = symbol, y = avg_predicted_price, fill = symbol)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    title = "Average Predicted Stock Prices for December 2024",
    x = "Stock Symbol",
    y = "Average Predicted Price",
    fill = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_brewer(palette = "Set2")

```

## Average Predicted Stock Price - December

```{r}
# Calculate average predicted price for December
dec_avg_prices <- dec_results %>%
  group_by(symbol) %>%
  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE)) %>%
  mutate(avg_predicted_price = dollar(round(avg_predicted_price, 2))) %>%  # Format as dollar and round
  arrange(desc(avg_predicted_price))  # Sort by average price in descending order

# Display the table
dec_avg_prices %>%
  kable(
    caption = "Average Predicted Stock Prices for December 2024",
    col.names = c("Stock Symbol", "Average Predicted Price"),
    format = "markdown"  # Output as markdown for readability
  )
```

## Combined November and December
```{r}
# Combine November and December results
combined_results <- bind_rows(
  nov_results %>% mutate(month = "November"),
  dec_results %>% mutate(month = "December")
)

# Plot predicted prices with facets for each stock
ggplot(combined_results, aes(x = as.Date(date), y = predicted_adjusted, color = month)) +
  geom_line(size = 1) +
  facet_wrap(~symbol, scales = "free_y") +
  labs(
    title = "Predicted Stock Prices for November and December 2024",
    x = "Date",
    y = "Predicted Stock Price",
    color = "Month"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

## Average Predicted Stock Prices for November and December 2024

```{r}
# Reorder the month factor so that "November" comes before "December"
avg_predicted_prices <- avg_predicted_prices %>%
  mutate(month = factor(month, levels = c("November", "December")))

# Plot with the updated order
ggplot(avg_predicted_prices, aes(x = symbol, y = mean_price, fill = month)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("November" = "#F8766D", "December" = "#00BFC4")) +
  labs(
    title = "Average Predicted Stock Prices for November and December 2024",
    x = "Stock Symbol",
    y = "Average Predicted Price (in USD)",
    fill = "Month"
  )
```
